### 2023-07-11 paper/paper.Rmd
writing learning results

---
title: "Learning from Data Through Models"
date: '`r format(Sys.time(), "%B %d, %Y")`'
author: "Alberto Bisin, Guillaume Frechette and Jimena Galindo"
abstract: "TBW"
output: 
    pdf_document:
        citation_package: natbib
        fig_width: 7
        fig_height: 6
        fig_caption: true
        number_sections: false
        keep_tex: true
        extra_dependencies:
            footmisc: ["bottom"] # footnote management
            setspace: ["doublespacing"] # spacing of the paper
            caption: ["normal"]
            dsfont: null # indicator function 1
            booktabs: null
            makecell: null
            hyperref: null

bibliography: bibliography.bib
fontsize: 12pt
geometry:
- margin=2.5cm
hyperref:
- linestretch=2.0
documentclass: article
---

```{r dependencies, include=FALSE}
library(data.table)
library(here)
renv::snapshot()
library("languageserver")
library(reticulate)
```

# Introduction
TBW

# Literature Review
TBW

# Framework and Predictions

A finite number of observable random variables, $X =(x_1, x_2, ..., x_N)$, determine a binary state, $y$. 
The observable variables are independently distributed and each follows a Normal distribution with corresponding mean 
$\mu_i$ and standard deviation $\sigma_i$ with $i \in \{1, 2, ..., N\}$.

The state is \emph{Red} (**R**) or \emph{Blue} (**B**) and is determined in the following way:
\begin{equation*}
y = \begin{cases}
\text{\textbf{B}} &\text{if $a_1x_1+...+a_Nx_N + K\geq0$}\\
\text{\textbf{R}} &\text{otherwise}
\end{cases}
\end{equation*}

An observable variable $x_i$ is *relevant* if the associated coefficient, $a_i$, is non-zero. 
Likewise, variable $i$ is *irrelevant* if $a_i = 0$. 
We call the set of relevant variables $\mathcal{R}$ and the set of irrelevant variables $\mathcal{I}$.

In each period, the decision-maker (DM) observes the realization of all observable variables and predicts the state. 
Their flow payoff in period $t$ is 1 if they predict the state correctly, and it is zero otherwise. 
Denote the prediction by $\hat{y}$.

In order to make a prediction, the DM uses a map $f: M_t\to \{\textbf{B}, \textbf{R}\}$  
where $M_t \subseteq \{x_1, x_2, ..., x_N\}$ is the set of variables that they choose to consider in period $t$. 
Our main object of study is the set of variables that the DM decides to consider and not the properties of the 
mapping from the variables to the state. We refer to the set of variables that are taken into consideration, $M_t$, 
as the DM's *model of the world*, and we abstract away from how the variables are mapped into a prediction of 
the state.

A model of the world is said to be *simple* if it does not include all the relevant variables, 
*i.e.* if $M\subset \mathcal{R}$. And it is complex if it includes all relevant variables, $M=\mathcal{R}$. Thus, whenever 
agents are using simple models, they are ignoring some of the relevant variables even if the information is available to them.
In contrast, whenever agents are using complex models, they are using all the relevant information that is available to them, 
so restricting the number of variables that is available to them will impair their ability to predict correctly. In the design
of the experiment we will exploit these differences in order to determine whether subjects are using simple or complex models.

**Prediction 1:** If subjects are using simple model with $K$ variables, then allowing them to disclose $K+1$ variables 
will not increase the frequency with which they make correct predictions.

**Prediction 1B:** If subjects are using complex models, then allowing them to disclose an additional variable will always 
increase the frequency with which they make correct predictions.

### Optimal Rules
The most accurate prediction rule for this setting will be one that predicts the state to be **B** whenever, 
given the observables, **B** is more likely. And it predicts **R** whenever **R** is more likely 
given the observables. Because we abstract away from the procedure through which the rule is updated, 
we will make predictions under the assumption that the DM has learned the parameters perfectly. That is, 
they have seen enough information that, under an updating procedure that converges to the truth---for example, 
correctly specified regression model or Bayes rule---they would have converged to the true parameters already.

We refer to a prediction rule as an *Optimal Rule* if it is the rule that a fully rational Decision maker would 
converge to with perfect Bayesian learning. In this context, Bayesian learning corresponds to using Bayes rule to 
learn the coefficients $a = (a_1, a_2, ..., a_N)$ as well as the constant $K$. This means that in an $N$ dimensional 
space, the Decision Maker must learn $N+1$ parameters which corresponds to learning the halfspace in which the state 
is **B**. The optimal rule for the complex model coincides with the true data-generating process when learning 
is perfect.

Conditional on using a simple model, the optimal rule for that model is the rule that, when restricted to the 
variables being considered, does the best at predicting the state. In this case, the Decision Maker is learning only 
the parameters that pertain the variables in the model. Hence, for a model that considers $K$ variables, the DM will 
have to learn $K+1$ parameters. The following proposition illustrates exactly what the optimal prediction rules look 
like when the DM is able to learn.

##### Proposition.
_Fix any model that considers $k\leq N$ variables and relabel the variable so that $x_1, ..., x_K$ are the variables 
that the model considers and $x_{k+1}, ..., x_N$ are the ones that the model does not consider. Relabel the 
coefficients $a = (a_1, ..., a_N)$ accordingly as well._

##### proof.
_Let $M:=a_1x_1+...+a_kx_k$ and $M^c = a_{k+1}x_{k+1}+...+a_Nx_N+k$ and define the latent variable $y^L := M +  M^c$. 
The optimal prediction rule for a model that considers the variables in $M$ the state to be $\textbf{B}$ whenever 
$P[y^L|M]\geq \frac{1}{2}$ and predicts $\textbf{R}$ otherwise._

_Using the fact that for a random variable $z\sim N(\mu_z, \sigma_z)$ we have that $P[z\geq \mu_z]=\frac{1}{2}$ and 
noticing that $M^c$ and $y^L|M$ are Normally distributed with means $\mathbb{E}[M^c]$ and $M+\mathbb{E}[M^c]$ 
respectively, it is easy to see that  $P[y^L\geq 0 | M] \geq P[y^L\geq M+\mathbb{E}[M^c | M]]=\frac{1}{2}$ whenever 
$M+\mathbb{E}[M^c]\geq 0$. Similarly, whenever $M+\mathbb{E}[M^c]< 0$ we will have that 
$P[y^L\geq 0 | M] < P[y^L\geq M+\mathbb{E}[M^c | M]]=\frac{1}{2}$. Therefore, the optimal rule is to predict 
$\mathbf{B}$ whenever $a_1x_1+...+a_kx_k\geq -\mathbb{E}[M^c]$ and $\textbf{R}$ otherwise._

With limited data, it is not guaranteed that the DM is able to perfectly learn the parameters required for the optimal rule. 
However, the notion of optimal rules is useful for determining the scenarios in which we can expect polarization to arise. 
The next section uses the notion of optimal rules to formalize the concept of polarization and to establish the 
main prediction of the theory.

### Polarization
To study polarization, we take the definition of @Haghtalab2021: two DMs are said to be 
polarized if, when observing the same realization of observables, they predict the state to be different. That is, 
they both observe the same draw of $x=(x_1, ..., x_N)$ but one DM predicts the state to be **B** and the other 
predicts the state to be **R**.

Within out framework---as in @Haghtalab2021---polarization may persist even when DMs have 
access to infinite data because they might be using different models of the world in order to make their predictions. 
Nevertheless using different models does not necessarily imply that beliefs will always be polarized; it must also be 
the case that the realization of observables falls in a *disagreement zone*. By disagreement zone, we simply 
mean that if the two models that are being used for prediction are $M_1$ and $M_2$, the realization of the 
observables $x=(x_1, ..., x_N)$ must be such that $M_1$ makes a different prediction from $M_2$. The following 
2-dimensional example illustrates why the use of different models is necessary but not sufficient for polarization 
to arise.

##### Example
_The optimal rule for the complex model coincides with the truth, it predicts **B** if $x_2+1\geq x_1$. 
In addition, there are two simple models: one that considers only $x_1$---call it $M_1$---and one that considers 
only $x_2$---call it $M_2$. The optimal rules for $M_1$ and $M_2$ are threshold rules that predict the state is **B** 
if the value of $x_i$ is below (above) the threshold $t_i := \mu_{-i} + k$ and the predicted state is **R** otherwise.
These prediction rules are illustrated in . As the figure makes clear, both rules agree on the predicted state for 
some values of $(x_1, x_2)$ and disagree for other values. Theregions in which the rules make different predictions
are the disagreement zones._

For any pair of agents who observe the same realization of teh observables and use different models, it is possible
to determine whether they will be polarized or not by looking at the predictions made by their models. Notice that the 
polarizing observations---those that fall into a polarization zone---are different for each pair of models. Also notice that 
our definition of the polarization zone assumes that the DMs are using optimal rules. This is a very strong assumption 
and will most likely not be the case that subjects in the experiment learn perfectly. However, the notion of optimal rules
allows us to have a clean definition of polarization and to make predictions about when polarization will arise.

**Prediction 2 :** Polarization will arise more often when the theory predicts so. That is, when two agents who have 
different models of the world, are facing a realization of observables that falls into the polarization zone corresponding 
to their models.

# Experimental Design
```{r read-data-all}
data <- fread(here("data", "clean", "all.csv"))
part1 <- fread(here("data", "clean", "part1.csv"))
part2 <- fread(here("data", "clean", "part2.csv"))

```

Subjects for the experimet were recruited from the CESS lab subject pool at NYU. 
We recruited `r data[, uniqueN(participant.code)]` undergraduate students who participated in this experiment in person at 
the CESS lab. The experiment was coded using o-tree @otree and it consisted of two parts: Part 1 was meant to expose the 
subjects to the data generating process
and allow them to learn how to predict; part 2 was meant to elicit the models of the world that subjects might have
developed in part 1.

Upon arrival subjects were randomly placed in pairs, however, they were not informed of this fact.
Part 1 consited of 20 rounds in which subjects observed the realizations of all of the observables and were asked to predict
the state. We had 5 observable variables called variable 1, 2, 3, 4 and 5. The variables were independent and Normally 
distributed with means and variances as described in TABLE.

The state was determined by the following rule: $11x_1+6x_2+4.5x_3+2.4*x_4+k \geq 0$. The values of the coefficients were chosen
so that all variables had a similar but unique level of informativeness when considered on their own. In this case, because
all the variables are normally distributed, the information that they convey about the state is related to the
variance of the modified normal $a_ix_i$. The constant $k$ was chosen so that the state was **B** in 50% of the cases. 
The last column of TABLE shows the probability of the optimal prediction rule making a correct prediction when the model used
considers only each variable by itself. This is the measure that we use to determine the informativeness of each variable and
is reported in the last column of TABLE.

| Variable | Mean | Variance | Modified Variable | Modified Variance | Informativeness |
| -------- | ---- | -------- | ----------------- | ----------------- | --------------- |
| $x_1$    | 0    |   1      | $11x_1$           |         11        |        .65      |
| $x_2$    | -10  |   2      | $6x_2$            |         12        |        .67      |
| $x_3$    | 5    |   3      | $4.5x_3$          |        13.5       |        .69      |
| $x_4$    | -5   |   4      | $2.4*x_4$         |         12        |        .63      |
| $x_5$    | 100  |   5      | $0x_5$            |         0         |         0       |

In each round, subjects were asked to predict the state. They were given feedback on whether their prediction was correct or not.
Both subjects within a pair were shown exactly the same realization of the variables and had to predict the same state. This allows
us to determine whether the pair was polarized  or not for that particular observation.

Part 2 started immediatly after part 1 and has 70 rounds. In this part each subject was randomly allowed to use only a certain 
number of variables in order to make their prediction. Instead of showing them the realizations of all 5 variables, they were 
told that they can disclose up to $K$ variables, where $K$ was drawn at random from the set $\{1, 2, ..., 5\}$. K was redrawn
independently every round and across subjects. We use the random assignment of $K$ to investigate whether the subjects are using 
siumple or complex models. If allowing subjects to disclose an additional variable improves their prediction accuracy, 
then it must be that they are in fact using the additional information for prediciton. If, on the other hand, they do not perform 
better when they have access to more information, it must be that although they are disclosing an additional variable, 
they do not correctly incorporate the infomation into their prediciton rule. Thus, if we observe that the performance of our 
subjects does not improve when they are allowed to disclose $K+1$ variables, 
relative to their performance when they were allowed to disclose only $K$ variables, their model of the world must be of size $K$
or lower. This feature design allows us to test prediciton 1 and 1B and determine the number of variables that subjects are using.

In order to explore the polarization prediciton, we use the fact that subjects are in fixed pairs throughout the experiment.
Both subjects in a pair observe the same realization of all the variables in each round. Therefore, we can determine whether
they are polarized or not by looking at whether they make the same prediction or not. We can then determine whether the theory 
predicts that they should be polarized or not by looking at the polarization zone corresponding to the models that they are using.
If the draw of the variables falls into the polarization zone, then the theory predicts that they should be polarized. The models
that we use for this prediction are the ones that include the variables that they chose to reveal in that round. It could be that
although they are revealing certain variables, they are not using them for prediction, however, we do not have a good way to 
determine this. Instead, we use the variables that they reveal as a proxy for the variables that they are using for prediction and 
 determine the polarization zones for those models.

# Results

In this section we present the results from the experiment. We start by exploring the general learning patterns of the subjects
as well as the model choices. We then look at the effect of allowing subjects to disclose an additional variable and lastly we
look at the polarization results.

### Learning
In the first part of the experiment, subjects were asked to predict the state given the realization of all the variables.
Having access to all the information they could predict the state perfectly if they managed to learn the parameters of the
hyperplane that determines the state. Figure \ref{learningAll} shows the share of correct predictions across rounds in part 1. 
We see that subjects are able to learn how to predict the state at a rate that is higher than random. They initially predict
the state corectly about 50% of the time, which is consistent with random guessing. The share of correct guesses increases to 
`r 100*round(part1[subsession.round_number>15, mean(player.correct)], digits=2)`% by the end of part 1. And this last number is 
significantly higher than 50% 
(p-value = `r round(t.test(part1[subsession.round_number>10, player.correct], mu=0.5)$p.value, digits=4)`). 
They continue to learn even after the first 20 rounds when we look at the rounds in which subjects had access to all the information.
By the end of the experiment the rate of correct predictions is 
`r 100*round(part2[subsession.round_number>30, mean(player.correct)], digits=2)`% which is still significantly higher than 50%
(p-value = `r round(t.test(part2[subsession.round_number>30, player.correct], mu=0.5)$p.value, digits=4)`).

```{r learningAll, out.width='50%', fig.align='center', fig.cap='\\label{learningAll} Share of correct predictions across rounds'}
knitr::include_graphics(here("computed_objects", "figures", "learning_all.png"))
```

# Conclusion

# Appendix