### 2023-07-11 paper/paper.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
  12pt,
]{article}
\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[margin=2.5cm]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage[bottom]{footmisc}
\usepackage[doublespacing]{setspace}
\usepackage[normal]{caption}
\usepackage{dsfont}
\usepackage{booktabs}
\usepackage{makecell}
\usepackage{hyperref}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{plainnat}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Learning from Data Through Models},
  pdfauthor={Alberto Bisin, Guillaume Frechette and Jimena Galindo},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\title{Learning from Data Through Models}
\author{Alberto Bisin, Guillaume Frechette and Jimena Galindo}
\date{July 11, 2023}

\begin{document}
\maketitle
\begin{abstract}
TBW
\end{abstract}

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

TBW

\hypertarget{literature-review}{%
\section{Literature Review}\label{literature-review}}

TBW

\hypertarget{framework-and-predictions}{%
\section{Framework and Predictions}\label{framework-and-predictions}}

A finite number of observable random variables,
\(X =(x_1, x_2, ..., x_N)\), determine a binary state, \(y\). The
observable variables are independently distributed and each follows a
Normal distribution with corresponding mean \(\mu_i\) and standard
deviation \(\sigma_i\). The state is \emph{Red} (\textbf{R}) or
\emph{Blue} (\textbf{B}) and is determined in the following way:

\begin{equation*}
y = \begin{cases}
\text{\textbf{B}} &\text{if $a_1x_1+...+a_Nx_N + K\geq0$}\\
\text{\textbf{R}} &\text{otherwise}
\end{cases}
\end{equation*}

An observable variable \(x_i\) is \emph{relevant} if the associated
coefficient, \(a_i\), is non-zero. Likewise, variable \(i\) is
\emph{irrelevant} if \(a_i = 0\). We call the set of relevant variables
\(\mathcal{R}\) and the set of irrelevant variables \(\mathcal{I}\).

In each period, an agent observes the realization of all observable
variables and predicts the state. Their flow payoff in period \(t\) is 1
if they predict the state correctly, and it is zero otherwise. Denote
the prediction by \(\hat{y}\).

In order to make a prediction, the agent uses a map
\(M_t\to \{\textbf{B}, \textbf{R}\}\) where
\(M_t \subseteq \{x_1, x_2, ..., x_N\}\) is the set of variables that
they choose to consider in period \(t\). Our main object of study is the
set of variables that the agent decides to consider in a given period,
and the procedure by which the maping or the set are updated. We refer
to the set of variables that are taken into consideration, \(M_t\), as
the agent's \emph{mental model}, and assume that the mapping is linear.

A mental model is \emph{simple} if it does not include all the relevant
variables, \emph{i.e.} if \(M\subset \mathcal{R}\). And it is
\emph{complex} if it includes all relevant variables, \(M=\mathcal{R}\).
In addition, mental models can be ranked in terms of their simplicity by
the number of relevant variables that they consider.

Whenever agents use simple models, they ignore some of the relevant
information that is available to them. In contrast, whenever agents use
complex models, they use all the relevant information available. This
means that when agents with complex mental models are restricted in
terms of how many variables they can consider, their ability to predict
the state will be impaired. On the contrary, restricting the number of
variables that an agent with a simple model can use, does not
necessarily affect their ability to predict the state. In particular, if
they are considering fewer variables than what the restriciton allows,
their predictions should be as accurate as without the restriciton.
These two effects are captured in predictions 1 and 1B respectively.

\textbf{Prediction 1:} If subjects use complex models, then allowing
them to use an additional variable will always increase the frequency
with which they make correct predictions.

\textbf{Prediction 1B:} If subjects use a simple mental model that
considers \(K\) variables, allowing them to use \(K+1\) variables will
not increase the frequency with which they make correct predictions.

\hypertarget{optimal-rules}{%
\subsubsection{Optimal Rules}\label{optimal-rules}}

Conditional of using a particular mental model, the most accurate
prediction rule that the agent can use is one that predicts the state to
be \textbf{B} whenever, given the observables under consideration,
\textbf{B} is more likely. And it predicts \textbf{R} whenever
\textbf{R} is more likely. Because we abstract away from the procedure
through which the rule is updated, we will make predictions under the
assumption that the agent has learned the parameters perfectly. That is,
they have seen enough information that, under an updating procedure that
converges to the truth---for example, correctly specified regression
model or Bayes rule---they would have converged to the true parameters
already. This provides an upper bound for how well any model can
perform.

We refer to a prediction rule as an \emph{Optimal Rule} if it is the
rule that a fully rational agent would converge to with perfect Bayesian
learning. In this context, Bayesian learning corresponds to using Bayes
rule to learn the coefficients \(a = (a_1, a_2, ..., a_N)\) as well as
the constant \(K\). This means that in an \(N\) dimensional space, the
agent must learn \(N+1\) parameters which corresponds to learning the
halfspace in which the state is \textbf{B}. The optimal rule for the
complex model coincides with the true data-generating process when
learning is perfect and the parameters are identified.

An optimal rule conditional on model \(M\) is the rule that, when
restricted to the variables being considered, does the best at
predicting the state. In this case, the agent is learning only the
parameters that pertain the variables in \(M\) plus a constant term. The
following proposition illustrates exactly what the optimal prediction
rules look like when the agent is able to learn perfectly.

\hypertarget{proposition.}{%
\subparagraph{Proposition.}\label{proposition.}}

\emph{Fix any model that considers \(m\leq N\) variables and relabel the
variable so that \(x_1, ..., x_m\) are the variables that the model
considers and \(x_{m+1}, ..., x_N\) are the ones that the model does not
consider. Relabel the coefficients \(a = (a_1, ..., a_N)\) accordingly
as well.}

\hypertarget{proof.}{%
\subparagraph{proof.}\label{proof.}}

\emph{Let \(M:=a_1x_1+...+a_mx_m\) and
\(M^c = a_{m+1}x_{m+1}+...+a_Nx_N+k\) and define the latent variable
\(y^L := M + M^c\). The optimal prediction rule for model \(M\) predicts
the state to be \(\textbf{B}\) whenever \(P[y^L|M]\geq \frac{1}{2}\) and
predicts \(\textbf{R}\) otherwise.}

\emph{Using the fact that for a random variable
\(z\sim N(\mu_z, \sigma_z)\) we have that \(P[z\geq \mu_z]=\frac{1}{2}\)
and noticing that \(M^c\) and \(y^L|M\) are Normally distributed with
means \(\mathbb{E}[M^c]\) and \(M+\mathbb{E}[M^c]\) respectively, it is
easy to see that
\(P[y^L\geq 0 | M] \geq P[y^L\geq M+\mathbb{E}[M^c | M]]=\frac{1}{2}\)
whenever \(M+\mathbb{E}[M^c]\geq 0\). Similarly, whenever
\(M+\mathbb{E}[M^c]< 0\) we will have that
\(P[y^L\geq 0 | M] < P[y^L\geq M+\mathbb{E}[M^c | M]]=\frac{1}{2}\).
Therefore, the optimal rule is to predict \(\mathbf{B}\) whenever
\(a_1x_1+...+a_kx_k\geq -\mathbb{E}[M^c]\) and \(\textbf{R}\)
otherwise.}

With limited data, it is not guaranteed that the agent is able to
perfectly learn the parameters required for the optimal rule. However,
the notion of optimal rules is useful for determining a benchmark for
how well each model can perform, as well as the scenarios in which we
can expect polarization to arise. The next section uses the notion of
optimal rules to formalize the concept of polarization and to establish
the main prediction of the theory.

\hypertarget{polarization}{%
\subsubsection{Polarization}\label{polarization}}

We take the definition of polarization from \citet{Haghtalab2021}: two
agents are said to be polarized if, when observing the same realization
of observables, they predict the state to be different. That is, they
both observe the same draw of \(x=(x_1, ..., x_N)\) but one agent
predicts the state to be \textbf{B} and the other predicts the state to
be \textbf{R}.

Within out framework---as in \citet{Haghtalab2021}---polarization may
persist even when agents have access to infinite data because they might
be using different models of the world. Nevertheless using different
models does not necessarily imply that beliefs will always be polarized,
it must also be the case that the realization of observables falls in a
\emph{disagreement zone}. By disagreement zone, we simply mean that if
the two models that are being used for prediction are \(M_1\) and
\(M_2\), the realization of the observables \(x=(x_1, ..., x_N)\) must
be such that \(M_1\) makes a different prediction from \(M_2\). The
following 2-dimensional example illustrates why the use of different
models is necessary but not sufficient for polarization to arise.

\hypertarget{example}{%
\subparagraph{Example}\label{example}}

\emph{The optimal rule for the complex model coincides with the truth,
it predicts \textbf{B} if \(x_2+1\geq x_1\). In addition, there are two
simple models: one that considers only \(x_1\)---call it \(M_1\)---and
one that considers only \(x_2\)---call it \(M_2\). The optimal rules for
\(M_1\) and \(M_2\) are threshold rules that predict the state is
\textbf{B} if the value of \(x_i\) is below (above) the threshold
\(t_i := \mu_{-i} + k\) and the predicted state is \textbf{R} otherwise.
These prediction rules are illustrated in . As Figure \ref{example}
makes clear, both rules agree on the predicted state for some values of
\((x_1, x_2)\) and disagree for other values. The regions in which the
rules make different predictions are the disagreement zones.}

\begin{figure}

{\centering \includegraphics[width=0.5\linewidth]{../diagram_polariazation_2variables}

}

\caption{\label{example} Polarization regions for two single variable models}\label{fig:example}
\end{figure}

For any pair of agents who observe the same realization of the
observables and use different models, it is possible to determine
whether they will be polarized or not by looking at the predictions made
by the optimal rule corresponding to their models. Notice that the
polarizing observations---those that fall into a polarization zone---are
different for each pair of models. Also notice that our definition of
the polarization zone assumes that the agents are using optimal rules.
This is a very strong assumption and will most likely not be the case
that subjects in the experiment learn perfectly. This is a theoretical
construct that allows us to formalize Prediction 2.

\textbf{Prediction 2 :} Polarization will arise more often when the
theory predicts so. That is, when two agents who have different models
of the world, are facing a realization of observables that falls into
the polarization zone corresponding to their models.

\hypertarget{experimental-design}{%
\section{Experimental Design}\label{experimental-design}}

Subjects for the experimet were recruited from the CESS lab subject pool
at NYU. We recruited 30 undergraduate students who participated in this
experiment in person durinf the summer and fall of 2023. The experiment
was coded using o-tree \citet{otree} and it consisted of two parts: Part
1 was meant to expose the subjects to the data generating process and
allow them to learn how to predict; part 2 was meant to elicit the
models of the world that subjects might have developed in part 1.

Part 1 consited of 20 rounds in which subjects observed the realizations
of all of the observables and were asked to predict the state. We had 5
observable variables called variable 1, 2, 3, 4 and 5. The variables
were independent and Normally distributed with means and variances as
described in TABLE.

The state was determined by the following rule:
\(11x_1+6x_2+4.5x_3+2.4*x_4+k \geq 0\). The values of the coefficients
were chosen so that all variables had a similar but unique level of
informativeness when considered on their own. In this case, because all
the variables are normally distributed, the information that they convey
about the state is related to the variance of the modified normal
\(a_ix_i\). The constant \(k\) was chosen so that the state was
\textbf{B} in 50\% of the cases. The last column of TABLE shows the
probability of the optimal prediction rule making a correct prediction
when the model used considers only each variable by itself. This is the
measure that we use to determine the informativeness of each variable
and is reported in the last column of TABLE.

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1159}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.0580}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1159}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2464}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2464}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.2174}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Mean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Variance
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Modified Variable
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Modified Variance
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
Informativeness
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\(x_1\) & 0 & 1 & \(11x_1\) & 11 & .65 \\
\(x_2\) & -10 & 2 & \(6x_2\) & 12 & .67 \\
\(x_3\) & 5 & 3 & \(4.5x_3\) & 13.5 & .69 \\
\(x_4\) & -5 & 4 & \(2.4*x_4\) & 12 & .63 \\
\(x_5\) & 100 & 5 & \(0x_5\) & 0 & 0 \\
\end{longtable}

In each round, subjects were asked to predict the state. They were given
feedback on whether their prediction was correct or not. Both subjects
within a pair were shown exactly the same realization of the variables
and had to predict the same state. This allows us to determine whether
the pair was polarized or not for that particular observation.

Part 2 started immediatly after part 1 and has 70 rounds. In this part
each subject was randomly allowed to use only a certain number of
variables in order to make their prediction. Instead of showing them the
realizations of all 5 variables, they were told that they can disclose
up to \(K\) variables, where \(K\) was drawn at random from the set
\(\{1, 2, ..., 5\}\). K was redrawn independently every round and across
subjects. We use the random assignment of \(K\) to investigate whether
the subjects are using siumple or complex models. If allowing subjects
to disclose an additional variable improves their prediction accuracy,
then it must be that they are in fact using the additional information
for prediciton. If, on the other hand, they do not perform better when
they have access to more information, it must be that although they are
disclosing an additional variable, they do not correctly incorporate the
infomation into their prediciton rule. Thus, if we observe that the
performance of our subjects does not improve when they are allowed to
disclose \(K+1\) variables, relative to their performance when they were
allowed to disclose only \(K\) variables, their model of the world must
be of size \(K\) or lower. This feature design allows us to test
prediciton 1 and 1B and determine the number of variables that subjects
are using.

In order to explore the polarization prediciton, we use the fact that
subjects are in fixed pairs throughout the experiment. Both subjects in
a pair observe the same realization of all the variables in each round.
Therefore, we can determine whether they are polarized or not by looking
at whether they make the same prediction or not. We can then determine
whether the theory predicts that they should be polarized or not by
looking at the polarization zone corresponding to the models that they
are using. If the draw of the variables falls into the polarization
zone, then the theory predicts that they should be polarized. The models
that we use for this prediction are the ones that include the variables
that they chose to reveal in that round. It could be that although they
are revealing certain variables, they are not using them for prediction,
however, we do not have a good way to determine this. Instead, we use
the variables that they reveal as a proxy for the variables that they
are using for prediction and determine the polarization zones for those
models.

\hypertarget{results}{%
\section{Results}\label{results}}

In this section we present the results from the experiment. We start by
exploring the general learning patterns of the subjects as well as the
model choices. We then look at the effect of allowing subjects to
disclose an additional variable and lastly we look at the polarization
results.

\hypertarget{learning}{%
\subsubsection{Learning}\label{learning}}

In the first part of the experiment, subjects were asked to predict the
state given the realization of all the variables. Having access to all
the information they could predict the state perfectly if they managed
to learn the parameters of the hyperplane that determines the state.
Figure \ref{learningAll} shows the share of correct predictions across
rounds in part 1. We see that subjects are able to learn how to predict
the state at a rate that is higher than random. They initially predict
the state corectly about 50\% of the time, which is consistent with
random guessing. The share of correct guesses increases to 59\% by the
end of part 1. And this last number is significantly higher than 50\%
(p-value = \ensuremath{3\times 10^{-4}}). They continue to learn even
after the first 20 rounds when we look at the rounds in which subjects
had access to all the information. By the end of the experiment the rate
of correct predictions is 59\% which is still significantly higher than
50\% (p-value = 0).

\begin{figure}

{\centering \includegraphics[width=0.5\linewidth]{../computed_objects/figures/learning_all}

}

\caption{\label{learningAll} Share of correct predictions across rounds}\label{fig:learningAll}
\end{figure}

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\renewcommand\refname{Appendix}
  \bibliography{bibliography.bib}

\end{document}